---
title: "reserveSeedList"
author: "Melina Houle"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Intro

This script creates Builder input seed.csv and nghbr.csv files using for a reserve. The user can then run Builder externally and come back to the script to generate the upstream and downstream polygon based on the builder output file ROW_UPSTREAM_CATCHMENTS and ROW_DOWNSTREAM_CATCHMENTS. 

The first step is to locate catchments and protected areas shapefile representing the reserves and point on an output directory where all derived files will be saved. Catchments and protected areas shapefile must be using the same CRS. 

```{r Params, echo=TRUE}
# Load libraries
library(sf)
library(plyr)
library(dplyr)
library(utils)
library(stringr)
library(readr)
library(tibble)
source("./reserveSeedList_dependencies.R")
source("./dci.R")

#Set access path 
dirpath <- "E:/MelinaStuff/BEACONs/request/Kim/preBuilder"

#Set access path catchments layer
catchments <- file.path(dirpath ,"NCC_Quebec/debug_files_2025-03-19/input", "QC_catch.shp")

#Set access path protected areas layer
pas <- file.path(dirpath, "NCC_Quebec/debug_files_2025-03-19/input", "QC_pas.shp")
pas_colName <- "PAID"

#Set access path streams layer
streams <- file.path(dirpath, "NCC_Quebec/debug_files_2025-03-19/input", "streams.shp")

#Build folder trees for putput
treedir <- c("shp_output","Builder_input","Builder_output")
for(d in treedir){
  if(!dir.exists(file.path(dirpath, d))){
    dir.create(file.path(dirpath, d))
  }
}

# Initialize sf objects
catchments_sf <- st_read(catchments)
pas_sf <- st_read(pas)
streams_sf <- st_read(streams)
```

## 1. Generate reserve seedlist and neighbour input file 

The script initializes catchments and reserve and generates the Builder input from them. 

Function `genReserveSeed` need the following:
- **catchments**:  Point to a catchments sf object
- **reserves**: Point to a reserve sf object (protected areas)
- **reserveName**: Indicate the column name for the reserve unique identifier within the reserve sf object (i.e. `PA_ID`). Column containing special character can't be used
- **areaTarget**: Set an areaTarget
- **joinType**: Indicate how CATCHNUM intersecting a reserve will be retrieved. jointType can be INTERSECT or CENTROID
- **outDir**: Point to an output directory

```{r input, echo=TRUE}
# Create neighbour file
nghbrs <- neighbours(catchments_sf)
write.csv(nghbrs, file=file.path(dirpath,"Builder_input/nghbrs.csv"), row.names=FALSE) # Convert neighbours table to csv file.

#Create reserve seed
seed <- genReserveSeed(catchments_sf, pas_sf, reserveName = pas_colName, areaTarget = 10000, joinType = "INTERSECT", outDir = dirpath)

```

## 2. Run Builder externally 

In Builder, select seeds and nghb from the Builder_input folder and point to Builder_output folder for the output directory

## 3. Generate upstream and downstream polygons from catchments using Builder output files

Function `gen_poly_from_rowTable` allow to read Builder ouptut tables that uses rows format and generate polygons  using enumerated catchnum per reserve id. `gen_poly_from_rowTable` needs the following:

- **catchments**: Point to a catchments sf object
- **outDir**: Point to an output directory
- **type**: Indicate which file to read. type can be UPSTREAM or DOWNSTREAM

```{r genPoly, echo=TRUE}
#upstream
up_poly <- gen_poly_from_rowTable(catchments_sf, outDir = dirpath, type = "UPSTREAM")
write_sf(up_poly, dsn =file.path(dirpath, "shp_output", "out_up.shp"), append = FALSE)

#downstream
down_poly <- gen_poly_from_rowTable(catchments_sf, outDir = dirpath, type = "DOWNSTREAM")
write_sf(down_poly, dsn=file.path(dirpath, "shp_output","out_down.shp"), append = FALSE)
```


## 4. Calculate DCI

Function `calc_dci` calculate dendritic connectivity index (DCI) for a set of input polygons. Values range between 0 and 1, with 1 indicating a completely connected river network. Dendritic connectivity measures the longitudinal connectivity of a river network inside a set of reserves based on the expected probability of an organism being able to move between two random points in the network. 

Function `calc_lwdci` calculate stream network length-weighted dendritic connectivity index (DCI) for a set of input polygons. Dendritic connectivity measures the longitudinal connectivity of a river network inside a set of conservation areas based on the expected probability of an organism being able to move between two random points in the network. 

Both functions need the following:

- **conservation_area_sf**: Point to a protected areas sf object
- **stream_sf**: Point to an stream_sf object. The provided streams need to have information on BASIN. 
- **column_id**: Column name  holding the unique ID of the conservation_area_sf. Default is network. 
- **buffer_width**: Width of buffer to apply to stream segments. Defaults to 0.1. Used to ensure adjacent stream segments are connected during analysis.

```{r dci, echo=TRUE}
#DCI
pas_sf_with_dci <- pas_sf %>%
  mutate(dci = sapply(1:nrow(pas_sf), function(i) {
    pa <- pas_sf[i, ]
    calc_dci(pa, streams_sf, column_id = pas_colName)  # Assuming calc_dci is a function
  }))

pas_sf_with_lwdci <- pas_sf_with_dci %>%
  mutate(lwdci = sapply(1:nrow(pas_sf_with_dci), function(i) {
    pa <- pas_sf_with_dci[i, ]
    calc_lwdci(pa, streams_sf, column_id = pas_colName)  # Assuming calc_dci is a function
  }))

#Save results
write_sf(pas_sf_with_lwdci, dsn=file.path(dirpath, "shp_output","pas_with_dci.shp"), append = FALSE)
```