## BEACONs Representation R Functions

# `gen_targets()`

The `gen_targets()` takes a reference area polygon and a classified raster for which representation will be evaluated. The functiobn calculates the proportion of each class in the reference area and multiples by reserve size to get target in km2.


## Usage
```r
gen_targets(
  reference_sf, 
  representation_raster, 
  reserve_size, 
  class_values = c())
)

```
## Arguments
- reference_sf: sf object of the reference area we are aiming to represent.
- representation_raster: Raster object of the representation layer classified into categorical classes.
- reserve_size: A vector of classes in representation_raster to generate targets for. Defaults to all classes in the representation_raster.
- class_values: The area in km2 that targets will sum to. Generally the approximate reserve size of reserves being evaluated.



## 📤 Output

A tibble with columns: 
- `class_value`: the list of class_values  
- `area_km2`: area of the class value in the reference_sf  
- `reserve_size`: the provided reserve_size  
- `class_proportion`: area_km2/sum(area_km2) 
- `target_km2`: class_proportion * reserve_area


## Examples
### Running the examples


1. Download and unzip BEACONs R Tools 

2. Run the examples below.

```{r evaluate_targets_using_catchments, fig.width=10, fig.height=6,  message=FALSE, warning=FALSE}
# Load libraries
library(sf)
library(terra)
library(dplyr)
library(here)
library(RColorBrewer)
library(tidyr)

# --------------------------------------
# SET PARAMS  --------------------
# --------------------------------------
# Set working directory
dirpath <- here(".")

setwd(dirpath)
source("./R/representation.R")
source("./R/utils.R")

#Set access path 
catchments_sf <- st_read(file.path(dirpath, "data/catchments_sample.shp"), quiet = TRUE)
nalc <- rast(file.path(dirpath, "data/nalc_sample.tif"))
refarea_sf <- st_read(file.path(dirpath, "data/reference_area_sample.shp"), quiet = TRUE)

```

## Setting targets
We will use a target area of 100km2 for this analysis to match the approximate size of the `beaconsbuilder` conservation areas. The `gen_targets()` function calculates targets and lists them in the column `target_km2`, calculated as the proportion of each class across the raster, multiplied by the user provided `reserve_size`.

Target tables can be created manually as long as they contain the columns "class_value", "class_proportion" and "target_km2". If fixed instead of proportional targets are desired the "target_km2" values could be edited using for example `target_table$target_km2 <- 50`.

```{r gen_targets, fig.width=10, fig.height=6,  message=FALSE, warning=FALSE}


# Generate target table using study area boundary and LED map
# Use the reserve size matching the builder area target (100km2)
target_table <- gen_targets(
  reference_sf = refarea_sf, 
  representation_raster = nalc, 
  reserve_size = 100)

print(as.data.frame(target_table))
```

```
